---
title: "Setting up Git, GitHub, and Rstudio - Cheatsheet"
author: "Sarah Wright and Rob Baker"
---


# Setting up Rstudio to talk with git

## 1. Make sure you have all the software and accounts you will need

-   Download Git from the Software Center
-   Download R from the software Center
-   Download Rstudio from the software Center
-   Register an account at github.com

## 2. Configure Git

You really should only have to do this once! From within Rstudio, access the "Terminal" pain and run the following code:

```{bash, eval=FALSE}         
git config --global user.name 'your_github_username'
git config --global user.email 'email_assoc_with_your_github_account'
git config --global --list
git config --global init.defaultBranch main
```

## 3. Configure Git with Rstudio

If you get API rate limit errors or your PAT expires, you may need to generate a new one. You may end up doing this multiple times but can generally skip it unless you run into errors or need to reset your credentials:

From with Rstudio's "Console" pane, run the following code:

If you don't already have it, install usethis and credentials for R:

```{r, install_packages, eval=FALSE}         
install.packages(c("usethis", "credentials")
```

create a personal access token for authentication:

```{r, create_PAT, eval=FALSE}
usethis::create_github_token()
```

running create_github_token() should open up a web browser and go to github.com. You may be asked to verify your account, enter a password, or otherwise link to the github ecosystem. Check for pop-ups on your second-monitor, if you have one!

From github.com, enter a description for your Personal Access Token (PAT) and select any options you might want; typically the default options work well for everything.

Don't forget to copy the PAT to your clipboard - you'll need it for the next step.

From within R, set your personal access token:

```{r, set_pat, eval=FALSE} 
credentials::set_github_pat("YourPAT")
```

# 4. Restart R

You should be all set!

# References, Acknowledgements, Resources

A very helpful github gist:
https://gist.github.com/Z3tt/3dab3535007acf108391649766409421

Stack overflow solutions:
https://stackoverflow.com/questions/69305874/authenticating-to-github-from-rstudio


Happy Git with R (very comprehensive):
https://happygitwithr.com/

GitHub PAT documentation:
https://docs.github.com/en/rest/orgs/personal-access-tokens?apiVersion=2022-11-28
